The Radio Sky at Meter Wavelengths: $m$-Mode Analysis Imaging\\ with the Owens Valley Long Wavelength Array

### 2017-10-26 paper.tex
add the paper as it will be submitted
package the paper for submission to arxiv

@@

% alphabetical by last name
\author{Benjamin~R.~Barsdell} \harvard \nvidia
\author{Stephen~A.~Bourke}    \astrocaltech \chalmers
\author{M.~A.~Clark}          \nvidia
\author{Steven~W.~Ellingson}  \vt
\author{Jayce~Dowell}         \unm
\author{Hugh~Garsden}         \harvard
\author{Lincoln~J.~Greenhill} \harvard
\author{Jonathon~Kocz}        \astrocaltech
\author{T.~Joseph~W.~Lazio}   \jpl
\author{Danny~C.~Price}       \harvard \berkeley \swinburne
\author{Frank~K.~Schinzel}    \nrao \unm
\author{Gregory~B.~Taylor}    \unm
\author{Yuankun~Wang}         \astrocaltech
\author{David~P.~Woody}       \ovro

@@

\begin{abstract}
    A host of new low-frequency radio telescopes seek to measure the 21-cm transition of neutral
    hydrogen from the early universe.  These telescopes have the potential to directly probe star
    and galaxy formation at redshifts $20 \gtrsim z \gtrsim 7$, but are limited by the dynamic range
    they can achieve against foreground sources of low-frequency radio emission. Consequently, there
    is a growing demand for modern, high-fidelity maps of the sky at frequencies below 200 MHz for
    use in foreground modeling and removal.  We describe a new widefield imaging technique for
    drift-scanning interferometers, Tikhonov-regularized $m$-mode analysis imaging.  This technique
    constructs images of the entire sky in a single synthesis imaging step with exact treatment of
    widefield effects.  We describe how the CLEAN algorithm can be adapted to deconvolve maps
    generated by $m$-mode analysis imaging.  We demonstrate Tikhonov-regularized $m$-mode analysis
    imaging using the Owens Valley Long Wavelength Array (OVRO-LWA) by generating 8 new maps of the
    sky north of $\delta=-30^\circ$ with 15 arcmin angular resolution and at frequencies evenly
    spaced between 36.528~MHz and 73.152~MHz. These maps are a 10-fold improvement in angular
    resolution over existing full-sky maps at comparable frequencies, which have angular resolutions
    $\ge 2^\circ$. The thermal noise in each map is $\sim$800 mJy/beam -- comparable to the
    confusion limit. However, each map is constructed exclusively from interferometric observations
    and does not represent the globally averaged sky brightness. Therefore future improvements will
    incorporate total power radiometry, and additional bandwidth and integration time to improve the
    thermal noise.  The angular resolution can also be further improved with the planned expansion
    of the OVRO-LWA to 2.6~km baselines.  These maps serve as a first step on the path to the use of
    more sophisticated foreground filters in 21-cm cosmology incorporating the measured angular and
    frequency structure of all foreground contaminants.
\end{abstract}

@@

Just as for CMB experiments, foreground removal or suppression is an essential component of both
classes of 21-cm cosmology experiments. The brightness temperature of the galactic synchrotron
emission at high galactic latitudes is measured by \citet{2008AJ....136..641R} as
\begin{equation}
    T \sim 300\,{\rm K} \times \left(\frac{\nu}{150\,{\rm MHz}}\right)^{-2.5}\,.
\end{equation}
Therefore, experiments conservatively need to achieve 5 orders of dynamic range against this
foreground emission before the cosmological signal can be measured. Current foreground removal
methods (for example, \citealt{2012ApJ...756..165P}) rely on the assumption that the foreground
emission is spectrally smooth. The low-frequency radio sky is composed of several components:
galactic synchrotron emission, supernova remnants, radio galaxies, free-free emission and absorption
from \ion{H}{2} regions, and a confusing background of radio sources.  Ideally, a foreground removal
strategy should be informed by the measured spatial structure and frequency spectrum of all
foreground components. For instance, CMB experiments typically construct several maps at several
frequencies to enable component separation.  At low frequencies, this possibility is limited by the
availability of suitable high-fidelity sky maps on angular scales ranging from tens of degrees to
arcminutes.

>>>>

\newcommand{\google}{
    \affiliation{
        Google,
        340 Main St,
        Venice, CA 90291, USA
    }
}

@@

% alphabetical by last name
\author{Benjamin~R.~Barsdell} \harvard \nvidia
\author{Stephen~A.~Bourke}    \astrocaltech \chalmers
\author{M.~A.~Clark}          \nvidia
\author{Steven~W.~Ellingson}  \vt
\author{Jayce~Dowell}         \unm
\author{Hugh~Garsden}         \harvard
\author{Lincoln~J.~Greenhill} \harvard
\author{Jacob~M.~Hartman}     \google
\author{Jonathon~Kocz}        \astrocaltech
\author{T.~Joseph~W.~Lazio}   \jpl
\author{Danny~C.~Price}       \harvard \berkeley \swinburne
\author{Frank~K.~Schinzel}    \nrao \unm
\author{Gregory~B.~Taylor}    \unm
\author{Harish~K.~Vedantham}  \astrocaltech
\author{Yuankun~Wang}         \astrocaltech
\author{David~P.~Woody}       \ovro

@@

\begin{abstract}
    A host of new low-frequency radio telescopes seek to measure the 21-cm transition of neutral
    hydrogen from the early universe.  These telescopes have the potential to directly probe star
    and galaxy formation at redshifts $20 \gtrsim z \gtrsim 7$, but are limited by the dynamic range
    they can achieve against foreground sources of low-frequency radio emission. Consequently, there
    is a growing demand for modern, high-fidelity maps of the sky at frequencies below 200 MHz for
    use in foreground modeling and removal.  We describe a new widefield imaging technique for
    drift-scanning interferometers, Tikhonov-regularized $m$-mode analysis imaging.  This technique
    constructs images of the entire sky in a single synthesis imaging step with exact treatment of
    widefield effects.  We describe how the CLEAN algorithm can be adapted to deconvolve maps
    generated by $m$-mode analysis imaging. We demonstrate Tikhonov-regularized $m$-mode analysis
    imaging using the Owens Valley Long Wavelength Array (OVRO-LWA) by generating 8 new maps of the
    sky north of $\delta=-30^\circ$ with 15 arcmin angular resolution, at frequencies evenly spaced
    between 36.528~MHz and 73.152~MHz, and $\sim$800 mJy/beam thermal noise. These maps are a
    10-fold improvement in angular resolution over existing full-sky maps at comparable frequencies,
    which have angular resolutions $\ge 2^\circ$. Each map is constructed exclusively from
    interferometric observations and does not represent the globally averaged sky brightness. Future
    improvements will incorporate total power radiometry, improved thermal noise, and improved
    angular resolution -- due to the planned expansion of the OVRO-LWA to 2.6~km baselines.  These
    maps serve as a first step on the path to the use of more sophisticated foreground filters in
    21-cm cosmology incorporating the measured angular and frequency structure of all foreground
    contaminants.
\end{abstract}

@@

Just as for CMB experiments, foreground removal or suppression is an essential component of both
classes of 21-cm cosmology experiments. The brightness temperature of the galactic synchrotron
emission at high galactic latitudes is measured by \citet{2017MNRAS.464.4995M} as
\begin{equation}
    T \sim 300\,{\rm K} \times \left(\frac{\nu}{150\,{\rm MHz}}\right)^{-2.6}\,.
\end{equation}
Therefore, experiments conservatively need to achieve 5 orders of dynamic range against this
foreground emission before the cosmological signal can be measured. Current foreground removal
methods (for example, \citealt{2012ApJ...756..165P}) rely on the assumption that the foreground
emission is spectrally smooth. The low-frequency radio sky is composed of several components:
galactic synchrotron emission, supernova remnants, radio galaxies, free-free emission and absorption
from \ion{H}{2} regions, and a confusing background of radio sources.  Ideally, a foreground removal
strategy should be informed by the measured spatial structure and frequency spectrum of all
foreground components. For instance, CMB experiments typically construct several maps at several
frequencies to enable component separation.  At low frequencies, this possibility is limited by the
availability of suitable high-fidelity sky maps on angular scales ranging from tens of degrees to
arcminutes.
